---
title: "ESM 206 Lab 5"
author: "Kat Leigh"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objectives:
- count data!
- use lubridate package to parse dates
- find confidence intervals and do T tests w/ t.test()
- Make 1st heatmap w/ geom_tile()

```{r, include = FALSE}
library(tidyverse)
library(here)
library(janitor)
```

```{r}
# import lobster data from subfolder
lobster_ab <- read_csv(here("data","lobster_abundance.csv"), 
  na = "-99999") %>% 
  clean_names()
```

(in tidyverse, every observation has its own role (every lobster). Currently it's in frequency format, and we need case format.)
```{r}
lobster_tidy <- lobster_ab %>% 
  uncount(lobster_count)
```
now each lobster has its own row, now it's in tidy format!
time to look at data
```{r}
ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) + 
  geom_jitter(aes(color = site),
              width = 0.2,
              alpha = 0.4)

ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) + 
  geom_violin(aes(color = site),
              alpha = 0.4)

ggplot(data= lobster_tidy, aes(x = size_mm))+
  geom_histogram(aes(fill = site))+
  facet_wrap(~site, scales = "free") #since we wanna see norm dist, change the scales, but usually don't do this if you're comparing ACROSS groups

ggplot(data= lobster_tidy, aes(sample = size_mm))+
  geom_qq(size = 0.7)+
  facet_wrap(~site, scales = "free")
```
convert data
```{r}
lobster_data <- lobster_tidy %>% 
  mutate(date_new = lubridate::mdy(date))
```

parse date column to get column that has month in it. and the column already is treated as a factor w/ assigned order!

```{r}
lobster_parse_date <- lobster_data %>% 
  mutate(obs_month = lubridate::month(date_new, label = TRUE)) %>% 
  mutate(obs_year = lubridate::year(date_new))
```
count lobsters by different groupings

counting number of lobsters by yr and month

```{r}
lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)
```
counting number of lobsters by sites

```{r}
lobster_ysite <- lobster_parse_date %>% 
  count(obs_year, site)
```
over all obs, how many lobsters at each site

```{r}
lobster_site_only <- lobster_parse_date %>% 
  count(site)
```

stick it all in 1 table

```{r}
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    sample_n = n()
  )
```

Confidence Intervals and t-tests
use ttest fxn to find confidence intervals and perform ttests

```{r}
ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)  #store a vector of values

t.test(ivee_lobsters) #one vector, so one sample t-test (is there anough evidence that the actual value is different from zero)
```

2 sample t-test!
Is there a sig dif btw mean lobs lengths btw naples reef and mohawk reef?

```{r}
napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(napl_sample,mohk_sample)

mohk_napl_ttest
```

(in help documentation, the value section refers to all the outputs of a fxn. use them to call/reference them!)

mean lobster size differed significatly btw Mohawk and Naples reefs(t(`r mohk_napl_ttest$parameter`) = `r mohk_napl_ttest$statistic`)

```{r}
lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL","MOHK"))

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)
ttest_2
```

make a heat map!

```{r}
ggplot(data = lobster_ysite, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

